The polynomial representation of the type $A_{n - 1}$ rational Cherednik algebra in characteristic $p \mid n$

### 2015-05-17 formalwriteup.tex
section 1,2 edits

Given the symmetric group $S_n$, we let $\mathcal{S}$ be the set of reflections in $S_n$. Let $\h$ be the irreducible $(n-1)$-dimensional representation of $S_n$, realized as the submodule of $k^n$ where the sum of the coordinates is $0$. For each $s \in \mathcal{S}$ we assign a vector $\alpha_s \in \h^*$ spanning the image of $1-s$, and choose $\alpha_s^\vee \in \h$ so that $(1-s)x=\langle \alpha_s^\vee,x\rangle \alpha_s$ for all $x \in \h^*$, where $\langle \cdot,\cdot\rangle$ indicates the pairing between $\h$ and $\h^*$.

Let $T(\h \oplus \h^*)$ be the tensor algebra of $\h \oplus \h^*$. We choose $\hbar,c \in k$. Then we define the {\it rational Cherednik algebra of type $A$} $\HH_{\hbar,c}(\h)$ as the quotient of $k[S_n] \ltimes T(\h \oplus \h^*)$ by the relations
\begin{align*}
[x,x']=0, \quad [y,y' = 0], \quad [y,x] = \hbar\langle y,x\rangle - \sum_{s \in \mathcal{S}} c\langle y,\alpha_s\rangle\langle \alpha_s^\vee,x\rangle s
\end{align*}
 for all $x,x' \in \h^*, y,y' \in \h$. We can give $\HH_{\hbar,c}(\h)$ a $\mathbb{Z}$-grading by setting $\deg x=1$ for $x \in \h^*$, $\deg y = -1$ for $y \in \h$, and $\deg g=0$ for $g \in k[G]$. We get the PBW-type decomposition $\HH_{\hbar,c}(\h)=\Sym(\h) \otimes_k k[S_n] \otimes_k \Sym(\h^*)$ (\cite[Section 3.2]{EM}).

\subsection{Representations of rational Cherednik algebras}

@@

As described in \cite[Section 2.5]{BC1}, $A=\Sym(\h^*)$ has a unique maximal graded proper submodule $J$ which can be realized as the kernel of the contravariant form $\beta_c: \Sym(\h^*) \times \Sym(\h) \to k$; $\beta_c$ can be characterized by the property that for all $x \in \h^*, y \in \h, f \in \Sym(\h^*), g \in  \Sym(\h)$:
\begin{align*}
\beta_c(xf,g)=\beta_c(f,xg), \quad \beta_c(f,yg) = \beta_c(yf,g), \quad \beta_c(1,1) = 1.
\end{align*}
The quotient $A/J$ is a finite-dimensional irreducible $\mathbb{Z}$-graded representation of $\HH_{\hbar,c}(\h)$. The {\it Hilbert series} $h_{A/J}(t)$ of $A/J$ is $\sum_{j=0}^\infty (\dim_k L_j)t^k$ where $L_j$ is the $j$-graded factor of $A/J$.

@@

 We call $c$ {\it generic} if we do not specify its value. We will be concerned with the case $\hbar=1$, $c$ generic, and the characteristic $p$ of $k$ dividing $n$. We realize $\h$ as the subspace of $k^n$ with the sum of the coordinates $0$, which is an irreducible representation of $S_n$ of dimension $n-1$. We state our result for this case below:

 \newtheorem*{thm:main}{Theorem \ref{thm:main}}
\begin{thm:main} The irreducible representation $A/J$  of $\HH_{\hbar,c}(\h)$ is a complete intersection with  Hilbert series $\left(\frac{1-t^p}{1-t}\right)^{n-1}$. The maximal ideal $J$ is generated by $[z^p]F_i(z)$ for the formal power series
\[
F_i(z)=\frac{1}{1-x_iz} \left( \sum_{k=0}^{p-1} \binom{c}{k}(-1+\prod_{j=0}^{n-1} (1-x_jz))^k\right) \text{ for $i=0,\dots,n-2$}.
\]
\end{thm:main}

%We assume $p>2$ since the case $p=2$ is fully characterized in \cite{L}.

@@

@@

To understand the action of $\HH_{\hbar,c}(\h)$ on $A$, we can use the PBW decompositions. The action of $\Sym(\h^*)$ on $A$ is by left multiplication; $k[S_n]$ acts by the diagonal action, and $\Sym(\h)$ acts via {\it Dunkl operators}. For $y \in \h$, the Dunkl operator $D_y$ acts on $A$ by:
\begin{align*}
D_y(f) = \hbar \partial_y f  - \sum_{s \in \mathcal{S}} c \frac{ ( y, \alpha_s )}{\alpha_s} (1-s). f.
\end{align*}
We choose bases for $\h,\h^*$ in the following way: let $V$ be the vector space spanned by $y_0,\dots,y_{n-1}$ and let $\h$ be the subspace spanned by $y_i-y_j$ for $i \ne j$; $S_n$ acts by permuting indices. Then if $x_0,\dots,x_{n-1}$ is the dual basis for $V^*$, we see that $\h^*$ is the span of $x_0,\dots,x_{n-1}$ under the relation $x_0+\dots+x_{n-1}=0$; alternatively we can consider $\h^*$ as the span of $x_0,\dots,x_{n-2}$ with $x_{n-1}$ defined as $-x_0-\dots-x_{n-2}$.  If $D_i$ is the Dunkl operator corresponding to $y_i$, the Dunkl operators for the elements of $\h$ are spanned by $D_i-D_j$ with $i \ne j$.

For a transposition $s_{ij} \in S_n$ with $i<j$, we let the corresponding vector $\alpha_{s_{ij}} \in \h^*$ be $x_i-x_j$.

@@

\begin{theorem}\label{thm:main} The irreducible representation $A/J$  of $\HH_{\hbar,c}(\h)$ is a complete intersection with  Hilbert series $\left(\frac{1-t^p}{1-t}\right)^{n-1}$. The maximal ideal $J$ is generated by $[z^p]F_i(z)$ for the formal power series
\[
F_i(z)=\frac{1}{1-x_iz} \left( \sum_{k=0}^{p-1} \binom{c}{k}(-1+\prod_{j=0}^{n-1} (1-x_jz))^k\right) \text{ for $i=0,\dots,n-2$}.
\]
\end{theorem}
\begin{proof}
It suffices to show that the $f_i$ generate the ideal $J$ and that $A/J$ has Hilbert series $\left(\frac{1-t^p}{1-t}\right)^{n-1}$.

>>>>

Given the symmetric group $S_n$, we see that $V=k^n$ is a representation of $S_n$ in a natural way (permutation of coordinates). Let $s_{ij}$ be the transposition permuting indices $i$ and $j$. We let $y_0,\dots,y_{n-1}$ be a basis of $V$; we define an $(n-1)$-dimensional representation $\h$ of $S_n$ as the span of $y_i-y_j$ for $i \ne j$. We choose a dual basis $x_0,\dots,x_{n-1}$ for $V^*$; the dual representation $\h^*$ of $\h$ can be realized as the quotient of $V^*$ by $x_0+\dots+x_{n-1}$. $S_n$ acts on $\h,\h^*$ by permuting the basis vectors in the natural way.

Let $T(\h \oplus \h^*)$ be the tensor algebra of $\h \oplus \h^*$. We choose $\hbar,c \in k$. Then we define the {\it rational Cherednik algebra of type $A$} $\HH_{\hbar,c}(\h)$ as the quotient of $k[S_n] \ltimes T(\h \oplus \h^*)$ by the relations
\begin{align*}
[x_i,x_j]=0, \quad [y_i,y_j] = 0, \quad [y_i,x_j] = \hbar\langle y_i,x_j\rangle - \sum_{k=0}^{n-1}\sum_{\ell=0}^{n-1} c\langle y_i,x_k-x_\ell \rangle\langle y_k-y_\ell,x_j\rangle s_{k\ell}
\end{align*}
 for all $0 \le i,j \le n-1$.

@@

@@

A statement holds for $c$ {\it generic} if it holds away from countably many values for $c$. Suppose that $\hbar=1$, $c$ generic, and the characteristic $p$ of $k$ dividing $n$.

 \newtheorem*{thm:main}{Theorem \ref{thm:main}}
\begin{thm:main} The irreducible quotient $A/J$ of the polynomial representation for $\HH_{\hbar,c}(\h)$ is a complete intersection with Hilbert series $\left(\frac{1-t^p}{1-t}\right)^{n-1}$. The maximal ideal $J$ is has linearly independent generators $f_i=[z^p]F_i(z)$ for $0 \le i \le n-2$ with
\[
F_i(z)=\frac{1}{1-x_iz} \left( \sum_{k=0}^{p-1} \binom{c}{k}(-1+\prod_{j=0}^{n-1} (1-x_jz))^k\right) \text{ for $i=0,\dots,n-2$}.
\]
\end{thm:main}

@@

\subsection{PBW decomposition}

We can give $\HH_{\hbar,c}(\h)$ a $\mathbb{Z}$-grading by setting $\deg x=1$ for $x \in \h^*$, $\deg y = -1$ for $y \in \h$, and $\deg g=0$ for $g \in k[G]$. We get the PBW-type decomposition $\HH_{\hbar,c}(\h)=\Sym(\h) \otimes_k k[S_n] \otimes_k \Sym(\h^*)$ (\cite[Section 3.2]{EM}).

As described in \cite[Section 2.5]{BC1}, $A=\Sym(\h^*)$ has a unique maximal graded proper submodule $J$ which can be realized as the kernel of the contravariant form $\beta_c: \Sym(\h^*) \times \Sym(\h) \to k$; $\beta_c$ can be characterized by the property that for all $x \in \h^*, y \in \h, f \in \Sym(\h^*), g \in  \Sym(\h)$:
\begin{align*}
\beta_c(xf,g)=\beta_c(f,xg), \quad \beta_c(f,yg) = \beta_c(yf,g), \quad \beta_c(1,1) = 1.
\end{align*}
The quotient $A/J$ is a finite-dimensional irreducible $\mathbb{Z}$-graded representation of $\HH_{\hbar,c}(\h)$. The {\it Hilbert series} $h_{A/J}(t)$ of $A/J$ is $\sum_{j=0}^\infty (\dim_k L_j)t^k$ where $L_j$ is the degree $j$ subspace of $L$.

@@

To understand the action of $\HH_{\hbar,c}(\h)$ on $A$, we can use the PBW decompositions. The action of $\Sym(\h^*)$ on $A$ is by left multiplication; $k[S_n]$ acts by the diagonal action, and $\Sym(\h)$ acts via {\it Dunkl operators}. For $y \in \h$, the Dunkl operator $D_y$ acts on $A$ by:
\begin{align*}
D_y(f) = \hbar \partial_y f  - \sum_{k=0}^{n-1} \sum_{\ell=0}^{n-1} c \frac{ \langle y, x_k-x_\ell \rangle}{x_k-x_\ell} (1-s_{k\ell}). f.
\end{align*}

 If $D_i$ is the Dunkl operator corresponding to $y_i$, the Dunkl operators for the elements of $\h$ are spanned by $D_i-D_j$ with $i \ne j$.

@@

\begin{theorem}\label{thm:main} The irreducible quotient $A/J$ of the polynomial representation for $\HH_{\hbar,c}(\h)$ is a complete intersection with Hilbert series $\left(\frac{1-t^p}{1-t}\right)^{n-1}$. The maximal ideal $J$ is has linearly independent generators $f_i=[z^p]F_i(z)$ for $0 \le i \le n-2$ with
\[
F_i(z)=\frac{1}{1-x_iz} \left( \sum_{k=0}^{p-1} \binom{c}{k}(-1+\prod_{j=0}^{n-1} (1-x_jz))^k\right) \text{ for $i=0,\dots,n-2$}.
\]
\end{theorem}
\begin{proof}
It suffices to show that the $f_i$ generate the ideal $J$ and that $A/J$ has Hilbert series $\left(\frac{1-t^p}{1-t}\right)^{n-1}$.
